Models including an interaction of treatment with baseline risk were better able to predict absolute benefit in simulations

### 2021-07-12 submission/manuscript.rmd
incorporate DvK comments

In the simulated datasets based treatment was allocated at random using a 50:50
split. For each patient we simulated $8$ baseline covariates, where
$x_1,\dots,x_4\sim N(0, 1)$ and $x_5,\dots,x_8\sim B(1, 0.2)$. Outcomes
for patients in the control arm were generated from a logistic regression model
including all baseline covariates. Coefficient values were such, so that the
prediction model had an AUC of $0.75$ and an event rate of $20\%$ in the control
arm was achieved. Under the base case scenario, outcomes in the treatment arm
were created using the same logistic regression model, including a constant
treatment odds ratio (OR) of $0.8$. The generated samples of the base case
scenario had sample size of 4,250 ($85%$ power for the detection of an
unadjusted OR of $0.8$).

We evaluated the effect of sample size considering additional scenarios with
sample sizes of `r 1064` and `r 17000`. We also evaluated the effect of prediction
performance, adjusting the baseline covariate coefficients, so that AUC values
of $0.65$ and $0.80$ were achieved when validating in a simulated dataset of
500,000 patients.

We simulated binary outcomes in the control arm using true probabilities 
$P(y=1|X) = \text{expit}(PI)$, where 
$\text{expit}(x)= \frac{e^x}{1 + e^x}$. 
In the treatment arm outcomes where generated with probabilities
$\text{expit}(lp_1)$ with
$$ lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0,$$
where the coefficients $\gamma_0, \gamma_1$ and $\gamma_2$ along with the
centering constant $c$ were set for each simulation scenario. In this way we
were able to assess a wide variety of scenarios, ranging from true constant
relative treatment effects to moderate and strong linear and quadratic
deviations. We also considered scenarios with treatment-covariate interactions.
These scenarios include 4 weak interactions
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.82$), 
4 strong interactions 
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61$), 
and 2 weak and 2 strong interactions. Combining all these different settings
resulted in a simulation study of $66$ scenarios. The exact settings for each
scenario are available in the supplementary material.

@@

All methods assume that a risk prediction model is available and can be used to
assign individualized risk predictions. For the simulations we developed the
prediction models internally and blinded to treatment using logistic regression
including main effects for all baseline covariates and treatment. Risk
predictions on individuals were made setting treatment to $0$.

The *stratified HTE method* has been suggested as an alternative to
traditional subgroup analyses. Patients are stratified into equally-sized risk
strata---in this case based on risk quartiles. Absolute effects are estimated
using the differences in event rates between treatments within risk quarters. We
considered this approach as a reference, expecting it to perform worse than the
other candidates, as its objective is not individualized benefit prediction.

Another approach would be to assume *constant relative treatment effect* (OR) is
constant. In that case, absolute benefit is estimated from
$\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))$. This is the method
usually considered when performing risk-based assessment of treatment effect
heterogeneity in RCT data.

A different approach fits a logistic regression using treatment, the prognostic
index and their interaction. In this case absolute benefit is estimated from
$\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)$.
We will refer to this method as the *linear interaction* approach.

Finally, we used *restricted cubic splines* (RCS) to relax the linearity assumption on
the effect of the linear predictor [@Harrell1988]. We compared the results
for 3, 4 and 5 knots when fitting the splines to introduce increasing
flexibility to the methods considered.

## Evaluation metrics

For evaluating the prediction error of the considered methods we used root mean
squared error (RMSE), since both the true and the predicted benefits are known,
given that this is a simulation study. More specifically, we calculated RMSE
from
$$\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2$$

We also compared the discriminative ability of the methods under study. We
assessed discrimination using the c-for-benefit statistic described in
[@vanKlaveren2018]. Patients in each treatment arm are ranked based on their
predicted benefit and then are matched 1:1, dropping patients in the larger
treatment arm without a pair. We define *observed* pair-specific treatment
benefit as the difference of observed outcomes between the untreated and the
treated patient of each pair. Pair-specific *predicted* benefit is defined as
the average of predicted benefits within each pair. Then, c-for-benefit is
defined as the probability that from two randomly chosen matched patient pairs
with unequal observed benefit, the pair with greater observed benefit also has a
higher predicted benefit.

We evaluated calibration in a similar manner, using the integrated calibration
index (ICI) for benefit [@Austin2019]. After creating pairs based on predicted
benefit, observed benefits are regressed on the predicted benefits using a
locally weighted scatterplot smoother (loess). The ICI is the average absolute
difference between predicted and smooth observed benefit. Values closer to $0$
represent better calibration.

@@

The model including a constant relative treatment effect had the lowest median
RMSE in scenarios with a true constant relative treatment effect (OR = 0.8, N =
`r 4250` and AUC = 0.75) or moderate relative deviations were considered (Figure
\ref{fig:rmsebase}; Panel A). However, when we considered strong linear and
quadratic deviations from a constant relative treatment effect the linear
interaction model performed best (Figure \ref{fig:rmsebase}; Panels B and C).
Only in the case of strong quadratic deviations models including RCS (3 knots)
performed equally well to the linear interaction method. Increasing the number
of knots in RCS resulted in higher error rates across all scenarios. The
adaptive approach performed very similarly to the best performing model in each
scenario.

@@

When we increased the sample size (N = `r 17000`), the model including a
constant relative treatment effect had the lowest RMSE under the assumption of
true constant relative treatment effects (Figure \ref{fig:rmsesamplesize};
Panel A). Contrary to previous results with smaller sample sizes, when
introducing moderate and strong linear deviations the linear interaction model
performed best, outperforming the constant treatment effect model (Figure
\ref{fig:rmsesamplesize}; Panels B and C). However, the linear interaction
model was outperformed by the more flexible RCS models (3 knots) in the case of
strong quadratic deviations. Again, the increased flexibility of RCS smoothing
with higher number of knots resulted in overfitting and worse performance
(Figure \ref{fig:rmsesamplesize}; Panel D).

@@

When we increased the true prediction AUC to 0.85, models including RCS
smoothing had the lowest RMSE in the presence of strong quadratic deviations
from the base case of true constant relative treatment effects (Figure
\ref{fig:rmseauc}; Panel D). However, with milder deviations, the linear
interaction model had the lowest RMSE with the RCS smoothing methods (3 knots)
being a close second (Figure \ref{fig:rmseauc}; Panels B and C). Increasing the
number of knots of RCS smoothing resulted in increased RMSE, which was less
pronounced in the case of strong quadratic deviations.

@@

The constant effects model, the linear interaction model and models with RCS
smoothing (3 knots) had the highest median c-for-benefit in the base case
scenario and the scenarios where linear and quadratic deviations were
considered. The constant treatment effect model and the linear interaction model
tended to present much lower variability compared to all other approaches
(Figure \ref{fig:discrimination}). We also observed an increasing trend of
discrimination for benefit variability and decreasing median values with
increasing number of restricted cubic spline knots in all scenarios.

@@

When focusing on calibration, the linear interaction model had the lowest median
ICI for benefit in the majority of scenarios except for the scenarios where no
or moderate linear deviations from the base case were considered. In that case
constant treatment effect models demonstrated the best performance, very
comparable to the linear interaction model's performance, nonetheless (Figure
\ref{fig:calibration}). However, under strong linear or
quadratic deviations, the constant treatment effect model was very poorly
calibrated (Figure \ref{fig:calibration}; Panels C and D).

@@

We used the risk linear predictor to fit the the proposed methods under study
for individualizing absolute benefit predictions. All methods had quite
comparable results, in the sense that we predicted increasing benefits for
patients with higher baseline risk predictions. In terms of c-for-benefit 
(validated internally) all models had quite comparable performance ranging from
0.519 (RCS smoothing with 3 knots) to 0.542 (RCS smoothing with 4 knots).
Similar conclusions could be drawn in terms of ICI-for-benefit which ranged
from 0.0039 (linear interaction approach) to 0.0053 (RCS smoothing with 4
knots).

The adaptive approach picked the
model with RCS smoothing with 4 knots, which had quite comparable performance to
the smooth fit with 5 knots. However, for lower baseline risk these 2 models
predicted implausible benefits and maybe should be avoided when applying such
models in practice. The linear interaction model and the model with RCS
smoothing (3 knots) made very similar predictions and also followed quite
closely the evolution of the stratified estimates. In this case, we observed
that even using a simple model with a constant relative treatment effect can
result in very similar benefit predictions as other more complex approaches.

>>>>

For each patient we generated 8 baseline covariates here
$x_1,\dots,x_4\sim N(0, 1)$ and $x_5,\dots,x_8\sim B(1, 0.2)$. Treatment
was allocated using a 50:50 split. Outcomes for patients in the control arm were
generated from a logistic regression model including all baseline covariates. In
the base case scenario coefficient values were such, that the AUC of the
logistic regression model was $0.75$ and the event rate in the control arm was
$20\%$ (Supplement 1).

While binary outcomes in the control arm were generated from Bernoulli variables
with true probabilities $P(y=1|X) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}$, 
the outcomes in the treatment arm based on true probabilities 
$\text{expit}(lp_1)$, with $$lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0$$

The coefficients $\gamma_0,\gamma_1$ and $\gamma_2$ along with the
centering constant $c$ were set for each simulation scenario to mimic a wide
variety scenarios, ranging from true constant relative treatment effect
($\gamma_1=\gamma_2=0$) to moderate and strong linear ($\gamma_2=0$) and
quadratic deviations. We also considered scenarios with treatment-covariate
interactions.
These scenarios include 4 weak interactions
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.82$), 
4 strong interactions 
($\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61$), 
and 2 weak and 2 strong interactions. Combining all these different settings
resulted in a simulation study of $66$ scenarios. The exact settings for each
scenario are available in the supplementary material.

The sample size of the base case scenario was set to 4,250 ($80%$ power for the
detection of a marginal OR of 0.8). We evaluated the effect of smaller or larger
sample sizes of 1,064 and 17,000, respectively. We also evaluated the of worse
or better discriminative ability for risk, adjusting the baseline covariate
coefficients, such that the AUC of the regression model in the control arm
was 0.65 and 0.85 respectively.

Combining all these settings resulted in a simualation study of 66 scenarios
(exact settings in the supplementary material).

@@

All methods assume that a risk prediction model is available to assign risk
predictions to individual patients. For the simulations we developed a
prediction model internally, using logistic regression including main effects
for all baseline covariates and treatment assignment. Risk predictions for
individual patients were based on treatment assignment to the control arm, that
is setting treatment assignment to 0.

A *stratified HTE method* has been suggested as an alternative to traditional
subgroup analyses. Patients are stratified into equally-sized risk strata---in
this case based on risk quartiles. Absolute treatment effects within risk strata
are estimated by the difference in event rate between patients in the control
arm and patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its objective is not
to individualize benefit prediction.

Second, we considered a model which assumes *constant relative treatment effect* (constant OR)
Hence, absolute benefit is predicted from
$\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))$.

Third, we considered a logistic regression model including treatment, the prognostic index, and their interaction. Absolute benefit is then estimated from 
$\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)$.
We will refer to this method as the *linear interaction* approach.

Finally, we used *restricted cubic splines* (RCS) to relax the linearity
assumption on the effect of the linear predictor [@Harrell1988]. We considered
splines with 3, 4 and 5 knots to compare models with different levels of
flexibility.

## Evaluation metrics
We evaluated the predictive accuracy of the considered methods by the root mean
squared error (RMSE):

$$\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2$$
We compared the discriminative ability of the methods under study using
c-for-benefit [@vanKlaveren2018]. The c-for-benefit represents the probability
that from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher predicted
benefit. To be able to calculate observed benefit, patients in each treatment
arm are ranked based on their predicted benefit and then matched 1:1 across
treatment arms. *Observed* treatment benefit is defined as the difference of
observed outcomes between the untreated and the treated patient of each matched
patient pair. *Predicted* benefit is defined as the average of predicted benefit
within each matched patient pair.

We evaluated calibration in a similar manner, using the integrated calibration
index (ICI) for benefit [@Austin2019]. The observed benefits are regressed on
the predicted benefits using a locally weighted scatterplot smoother (loess).
The ICI-for-benefit is the average absolute difference between predicted and
smooth observed benefit. Values closer to $0$ represent better calibration.

@@

The model including a constant relative treatment effect had the lowest median
RMSE in scenarios with a true constant relative treatment effect (OR = 0.8, N =
`r 4250` and AUC = 0.75) or with moderate relative deviations (Figure
\ref{fig:rmsebase}; Panel A). However, when we considered strong linear and
quadratic deviations from a constant relative treatment effect the linear
interaction model performed best (Figure \ref{fig:rmsebase}; Panels B and C).
Only in the case of strong quadratic deviations models including RCS (3 knots)
performed equally well to the linear interaction method. Increasing the number
of knots in RCS resulted in higher RMSE across all scenarios. The adaptive
approach performed very similarly to the best performing model in each scenario.

@@

When we increased the sample size (N = `r 17000`), the model including a
constant relative treatment effect had the lowest RMSE under the assumption of
true constant relative treatment effects (Figure \ref{fig:rmsesamplesize};
Panel A). However, when introducing moderate and strong linear deviations the
linear interaction model outperformed the constant treatment
effect model (Figure \ref{fig:rmsesamplesize}; Panels B and C). Furthermore, the
linear interaction model was outperformed by the more flexible RCS models (3
knots) in the case of strong quadratic deviations. Again, the increased
flexibility of RCS smoothing with higher number of knots resulted in overfitting
and worse predictive accuracy (Figure \ref{fig:rmsesamplesize}; Panel D).

@@

When we increased the AUC of the true prediction model to 0.85, models including
RCS smoothing had the lowest RMSE in the presence of strong quadratic deviations
from the base case of true constant relative treatment effects (Figure
\ref{fig:rmseauc}; Panel D). However, with milder deviations, the linear
interaction model had the lowest RMSE with the RCS smoothing methods (3 knots)
being a close second (Figure \ref{fig:rmseauc}; Panels B and C). Increasing the
number of knots of RCS smoothing resulted in increased RMSE, which was less
pronounced in the case of strong quadratic deviations.

@@

The constant effects model, the linear interaction model and models with RCS
smoothing (3 knots) had the highest median c-for-benefit in the base case
scenario and the scenarios where linear and quadratic deviations were
considered. The constant treatment effect model and the linear interaction model
tended to present much lower variability compared to all other approaches
(Figure \ref{fig:discrimination}). With increasing number of RCS knots, we
observed decreasing median values and increasing variability of the
c-for-benefit in all scenarios.

@@

As for calibration, the linear interaction model had the lowest median
ICI-for-benefit in the majority of scenarios except for the scenarios where no
or moderate linear deviations from the base case were considered. In that case
constant treatment effect models demonstrated optimal calibration, very
comparable to the linear interaction model's performance, nonetheless (Figure
\ref{fig:calibration}). With strong linear or quadratic deviations, the
constant treatment effect model was poorly calibrated (Figure
\ref{fig:calibration}; Panels C and D).

@@

We used the risk linear predictor to fit the proposed methods under study for
individualizing absolute benefit predictions. All methods predicted increasing
benefits for patients with higher baseline risk predictions, but the fitted
patterns were clearly different. The adaptive approach selected the model with
RCS smoothing with 4 knots. However, for very low baseline risk this model
predicted decreasing benefit with increasing risk maybe somewhat too flexible.
The more robust models, the linear interaction model or the model with RCS
smoothing (3 knots), gave very similar benefit predictions, followed the
evolution of the stratified estimates very closely and may therefore be
preferable for use in clinical practice.

### 2021-07-12 submission/manuscript.tex
incorporate DvK comments

In the simulated datasets based treatment was allocated at random using
a 50:50 split. For each patient we simulated \(8\) baseline covariates,
where \(x_1,\dots,x_4\sim N(0, 1)\) and \(x_5,\dots,x_8\sim B(1, 0.2)\).
Outcomes for patients in the control arm were generated from a logistic
regression model including all baseline covariates. Coefficient values
were such, so that the prediction model had an AUC of \(0.75\) and an
event rate of \(20\%\) in the control arm was achieved. Under the base
case scenario, outcomes in the treatment arm were created using the same
logistic regression model, including a constant treatment odds ratio
(OR) of \(0.8\). The generated samples of the base case scenario had
sample size of 4,250 (\(85%
\) power for the detection of an unadjusted OR of \(0.8\)).

We evaluated the effect of sample size considering additional scenarios
with sample sizes of 1,064 and 17,000. We also evaluated the effect of
prediction performance, adjusting the baseline covariate coefficients,
so that AUC values of \(0.65\) and \(0.80\) were achieved when
validating in a simulated dataset of 500,000 patients.

We simulated binary outcomes in the control arm using true probabilities
\(P(y=1|X) = \text{expit}(PI)\), where
\(\text{expit}(x)= \frac{e^x}{1 + e^x}\). In the treatment arm outcomes
where generated with probabilities \(\text{expit}(lp_1)\) with
\[ lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0,\] where the
coefficients \(\gamma_0, \gamma_1\) and \(\gamma_2\) along with the
centering constant \(c\) were set for each simulation scenario. In this
way we were able to assess a wide variety of scenarios, ranging from
true constant relative treatment effects to moderate and strong linear
and quadratic deviations. We also considered scenarios with
treatment-covariate interactions. These scenarios include 4 weak
interactions (\(\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.82\)), 4 strong
interactions (\(\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61\)), and 2
weak and 2 strong interactions. Combining all these different settings
resulted in a simulation study of \(66\) scenarios. The exact settings
for each scenario are available in the supplementary material.

@@

All methods assume that a risk prediction model is available and can be
used to assign individualized risk predictions. For the simulations we
developed the prediction models internally and blinded to treatment
using logistic regression including main effects for all baseline
covariates and treatment. Risk predictions on individuals were made
setting treatment to \(0\).

The \emph{stratified HTE method} has been suggested as an alternative to
traditional subgroup analyses. Patients are stratified into
equally-sized risk strata---in this case based on risk quartiles.
Absolute effects are estimated using the differences in event rates
between treatments within risk quarters. We considered this approach as
a reference, expecting it to perform worse than the other candidates, as
its objective is not individualized benefit prediction.

Another approach would be to assume \emph{constant relative treatment
effect} (OR) is constant. In that case, absolute benefit is estimated
from \(\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))\). This is
the method usually considered when performing risk-based assessment of
treatment effect heterogeneity in RCT data.

A different approach fits a logistic regression using treatment, the
prognostic index and their interaction. In this case absolute benefit is
estimated from
\(\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)\).
We will refer to this method as the \emph{linear interaction} approach.

Finally, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}7{]}. We
compared the results for 3, 4 and 5 knots when fitting the splines to
introduce increasing flexibility to the methods considered.

@@

For evaluating the prediction error of the considered methods we used
root mean squared error (RMSE), since both the true and the predicted
benefits are known, given that this is a simulation study. More
specifically, we calculated RMSE from
\[\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2\]

We also compared the discriminative ability of the methods under study.
We assessed discrimination using the c-for-benefit statistic described
in {[}8{]}. Patients in each treatment arm are ranked based on their
predicted benefit and then are matched 1:1, dropping patients in the
larger treatment arm without a pair. We define \emph{observed}
pair-specific treatment benefit as the difference of observed outcomes
between the untreated and the treated patient of each pair.
Pair-specific \emph{predicted} benefit is defined as the average of
predicted benefits within each pair. Then, c-for-benefit is defined as
the probability that from two randomly chosen matched patient pairs with
unequal observed benefit, the pair with greater observed benefit also
has a higher predicted benefit.

We evaluated calibration in a similar manner, using the integrated
calibration index (ICI) for benefit {[}9{]}. After creating pairs based
on predicted benefit, observed benefits are regressed on the predicted
benefits using a locally weighted scatterplot smoother (loess). The ICI
is the average absolute difference between predicted and smooth observed
benefit. Values closer to \(0\) represent better calibration.

@@

The model including a constant relative treatment effect had the lowest
median RMSE in scenarios with a true constant relative treatment effect
(OR = 0.8, N = 4,250 and AUC = 0.75) or moderate relative deviations
were considered (Figure \ref{fig:rmsebase}; Panel A). However, when we
considered strong linear and quadratic deviations from a constant
relative treatment effect the linear interaction model performed best
(Figure \ref{fig:rmsebase}; Panels B and C). Only in the case of strong
quadratic deviations models including RCS (3 knots) performed equally
well to the linear interaction method. Increasing the number of knots in
RCS resulted in higher error rates across all scenarios. The adaptive
approach performed very similarly to the best performing model in each
scenario.

@@

When we increased the sample size (N = 17,000), the model including a
constant relative treatment effect had the lowest RMSE under the
assumption of true constant relative treatment effects (Figure
\ref{fig:rmsesamplesize}; Panel A). Contrary to previous results with
smaller sample sizes, when introducing moderate and strong linear
deviations the linear interaction model performed best, outperforming
the constant treatment effect model (Figure \ref{fig:rmsesamplesize};
Panels B and C). However, the linear interaction model was outperformed
by the more flexible RCS models (3 knots) in the case of strong
quadratic deviations. Again, the increased flexibility of RCS smoothing
with higher number of knots resulted in overfitting and worse
performance (Figure \ref{fig:rmsesamplesize}; Panel D).

>>>>

For each patient we generated 8 baseline covariates here
\(x_1,\dots,x_4\sim N(0, 1)\) and \(x_5,\dots,x_8\sim B(1, 0.2)\).
Treatment was allocated using a 50:50 split. Outcomes for patients in
the control arm were generated from a logistic regression model
including all baseline covariates. In the base case scenario coefficient
values were such, that the AUC of the logistic regression model was
\(0.75\) and the event rate in the control arm was \(20\%\) (Supplement
1).

While binary outcomes in the control arm were generated from Bernoulli
variables with true probabilities
\(P(y=1|X) = \text{expit}(PI)=\frac{e^{PI}}{1+e^{PI}}\), the outcomes in
the treatment arm based on true probabilities \(\text{expit}(lp_1)\),
with \[lp_1 = \gamma_2(PI-c)^2 + \gamma_1(PI-c) + \gamma_0\]

The coefficients \(\gamma_0,\gamma_1\) and \(\gamma_2\) along with the
centering constant \(c\) were set for each simulation scenario to mimic
a wide variety scenarios, ranging from true constant relative treatment
effect (\(\gamma_1=\gamma_2=0\)) to moderate and strong linear
(\(\gamma_2=0\)) and quadratic deviations. We also considered scenarios
with treatment-covariate interactions. These scenarios include 4 weak
interactions (\(\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.82\)), 4 strong
interactions (\(\text{OR}_{t_x=1} / \text{OR}_{t_x=0}=0.61\)), and 2
weak and 2 strong interactions. Combining all these different settings
resulted in a simulation study of \(66\) scenarios. The exact settings
for each scenario are available in the supplementary material.

The sample size of the base case scenario was set to 4,250 (\(80%
\) power for the detection of a marginal OR of 0.8). We evaluated the
effect of smaller or larger sample sizes of 1,064 and 17,000,
respectively. We also evaluated the of worse or better discriminative
ability for risk, adjusting the baseline covariate coefficients, such
that the AUC of the regression model in the control arm was 0.65 and
0.85 respectively.

Combining all these settings resulted in a simualation study of 66
scenarios (exact settings in the supplementary material).

@@

All methods assume that a risk prediction model is available to assign
risk predictions to individual patients. For the simulations we
developed a prediction model internally, using logistic regression
including main effects for all baseline covariates and treatment
assignment. Risk predictions for individual patients were based on
treatment assignment to the control arm, that is setting treatment
assignment to 0.

A \emph{stratified HTE method} has been suggested as an alternative to
traditional subgroup analyses. Patients are stratified into
equally-sized risk strata---in this case based on risk quartiles.
Absolute treatment effects within risk strata are estimated by the
difference in event rate between patients in the control arm and
patients in the treated arm. We considered this approach as a reference,
expecting it to perform worse than the other candidates, as its
objective is not to individualize benefit prediction.

Second, we considered a model which assumes \emph{constant relative
treatment effect} (constant OR) Hence, absolute benefit is predicted
from \(\hat{\tau}(\bm{x}) = \text{expit}(PI +\log(\text{OR}))\).

Third, we considered a logistic regression model including treatment,
the prognostic index, and their interaction. Absolute benefit is then
estimated from
\(\hat{\tau}(\bm{x})=\text{expit}(\beta_0+\beta_{PI}PI) - \text{expit}(\beta_0+\beta_{t_x}+(\beta_{PI}+\beta_*)PI)\).
We will refer to this method as the \emph{linear interaction} approach.

Finally, we used \emph{restricted cubic splines} (RCS) to relax the
linearity assumption on the effect of the linear predictor {[}7{]}. We
considered splines with 3, 4 and 5 knots to compare models with
different levels of flexibility.

@@

We evaluated the predictive accuracy of the considered methods by the
root mean squared error (RMSE):

\[\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2\]
We compared the discriminative ability of the methods under study using
c-for-benefit {[}8{]}. The c-for-benefit represents the probability that
from two randomly chosen matched patient pairs with unequal observed
benefit, the pair with greater observed benefit also has a higher
predicted benefit. To be able to calculate observed benefit, patients in
each treatment arm are ranked based on their predicted benefit and then
matched 1:1 across treatment arms. \emph{Observed} treatment benefit is
defined as the difference of observed outcomes between the untreated and
the treated patient of each matched patient pair. \emph{Predicted}
benefit is defined as the average of predicted benefit within each
matched patient pair.

We evaluated calibration in a similar manner, using the integrated
calibration index (ICI) for benefit {[}9{]}. The observed benefits are
regressed on the predicted benefits using a locally weighted scatterplot
smoother (loess). The ICI-for-benefit is the average absolute difference
between predicted and smooth observed benefit. Values closer to \(0\)
represent better calibration.

@@

The model including a constant relative treatment effect had the lowest
median RMSE in scenarios with a true constant relative treatment effect
(OR = 0.8, N = 4,250 and AUC = 0.75) or with moderate relative
deviations (Figure \ref{fig:rmsebase}; Panel A). However, when we
considered strong linear and quadratic deviations from a constant
relative treatment effect the linear interaction model performed best
(Figure \ref{fig:rmsebase}; Panels B and C). Only in the case of strong
quadratic deviations models including RCS (3 knots) performed equally
well to the linear interaction method. Increasing the number of knots in
RCS resulted in higher RMSE across all scenarios. The adaptive approach
performed very similarly to the best performing model in each scenario.

@@

When we increased the sample size (N = 17,000), the model including a
constant relative treatment effect had the lowest RMSE under the
assumption of true constant relative treatment effects (Figure
\ref{fig:rmsesamplesize}; Panel A). However, when introducing moderate
and strong linear deviations the linear interaction model outperformed
the constant treatment effect model (Figure \ref{fig:rmsesamplesize};
Panels B and C). Furthermore, the linear interaction model was
outperformed by the more flexible RCS models (3 knots) in the case of
strong quadratic deviations. Again, the increased flexibility of RCS
smoothing with higher number of knots resulted in overfitting and worse
predictive accuracy (Figure \ref{fig:rmsesamplesize}; Panel D).
